---
title: "Untitled"
author: "Cyllia"
date: "2023-09-18"
output: html_document
---

##### Data
```{r}
AWTP <- read.csv("AWTP_MQE2023.csv")
awtp <- AWTP$Q1
```

##### Calculate Sample size
```{r}
n<-length(awtp)
n
```

##### Calculate Sample Mean
```{r}
Xbar <- mean(awtp)
Xbar
```

##### Calculate Sample Variance
```{r}
S2 <- var(awtp)
S2
```
##### Calculate Sample Standard Deviation
```{r}
S <- sd(awtp)
S
```
##### Calculate Standard Error of Xbar
```{r}
seXbar <- S/sqrt(n)
seXbar
```
###### Find a 95% confidence interval for the sample mean
##compute critical value for two-sided 95% confidence interval:
##need to use t-distribution with n-1 degrees of freedom, 
##so command qt for the .975 lowertail:
```{r}
cv <- qt(.975, df=n-1)   
cv
```
## if we just want to use 1-2-3 rule, we could just set cv <- 2
## if we know the population standard deviation set cv <- qnorm(.975)
#compute two-sided confidence interval *c(-1,1) for +/-:
```{r}
CI <- Xbar+cv*seXbar*c(-1,1)
CI
```
#alternatively, can use t.test command (but that is no fun):
```{r}
t.test(awtp, conf.level=0.95)
```
###### What is the probability that average awtp is larger than $15
```{r}
1-pnorm(15,Xbar,seXbar)
```
```{r}
p <- pnorm(3.05, 2.9, 0.2/5) - pnorm(2.95, 2.9, 0.2/5)
p
1-p

```
```{r}
mean(c(40, 55, 60, 90, 30, 50, 150, 100, 90, 80))
var(c(40, 55, 60, 90, 30, 50, 150, 100, 90, 80))
```

```{r}
data <- read.csv("Digital_Services_AWTP.csv")
data
```


